# Logônia Data Processing

## Overview

This document provides a step-by-step guide to download and process the [WorldClim 2.1](https://worldclim.org/) data **for the [Logônia](https://github.com/sustentarea/logonia) model**. For detailed instructions on how to download and process the data, please see the `LogoClim`'s [user manual](https://sustentarea.github.io/logoclim).

## Methods

### Source of Data

Data used in this notebook come from the following sources:

- [WorldClim](https://www.worldclim.org/)
  - [Historical Monthly Weather Data](https://www.worldclim.org/data/monthlywth.html) series, including 12 monthly data points for each year from 1951 to 2024. Based on [downscaled](https://worldclim.org/data/downscaling.html) data from [CRU-TS-4.09](https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.09/) (Harris et al., [2020](https://doi.org/10.1038/s41597-020-0453-3)), developed by the [Climatic Research Unit](https://www.uea.ac.uk/groups-and-centres/climatic-research-unit) at the [University of East Anglia](https://www.uea.ac.uk/). It provides monthly averages for minimum temperature, maximum temperature, and total precipitation.

### Data Munging

The data munging follows the data science workflow outlined by Wickham et al. ([2023](https://r4ds.hadley.nz/)). All processes were made using the [Quarto](https://quarto.org/) publishing system, the [R](https://www.r-project.org/) programming language, and several R packages.

For data manipulation and workflow, priority was given to packages from the [Tidyverse](https://www.tidyverse.org/), [rOpenSci](https://ropensci.org/) and [rspatial](https://rspatial.org/) frameworks, as well as other packages adhering to the [tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html).

### Code Style

The Tidyverse [tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html), [code style guide](https://style.tidyverse.org/) and [design principles](https://design.tidyverse.org/) were followed to ensure consistency and enhance readability.

### Reproducibility

The pipeline is fully reproducible and can be run again at any time. To ensure consistent results, the [`renv`](https://rstudio.github.io/renv/) package was used to manage and restore the R environment.

## Set Environment

```{r}
#| label: Set Environment
#| include: false

library(cli)
library(httr2)

if (!is_online()) {
  cli_abort(
    paste0(
      "No {.strong {col_red('internet connection')}} detected. ",
      "Please check your connection and try again."
    )
  )
}
```

### Load Packages

```{r}
#| code-fold: false
#| output: false

library(cli)
library(fs)
library(here)
library(httr2)
library(orbis) # github.com/sustentarea/orbis
```

### Set Data Directories

```{r}
data_dir <- here("data")
```

```{r}
if (!dir_exists(data_dir)) {
  dir_create(data_dir, recurse = TRUE)
}
```

### Set Parameters

:::{.callout-note}
The `box` variable defines the bounding box for the area of interest. It is specified in the format: `c(xmin, xmax, ymin, ymax)`. The coordinates are in decimal degrees (WGS84).

You can visualize the bounding box [here](http://bboxfinder.com/#-7.5,-63.5,0,-55).
:::

```{r}
box <- c(-63.5, -55, -7.5, 0)
```

::: {.callout-note}
Setting `ZIP_PROGRESS` environment variable is required by the [`zip`](https://r-lib.github.io/zip/) R package to display progress bars when unzipping data.
:::

```{r}
Sys.setenv(ZIP_PROGRESS = "true")
```

## Download Data

:::{.callout-note}
Keep in mind that the data can take up quite a bit of storage space. Make sure you have enough room before downloading.
:::

```{r}
files <- worldclim_download(
  series = "historical-monthly-weather-data",
  resolution = "10m",
  dir = data_dir
)
```

## Process Data

```{r}
files |> worldclim_to_ascii(box = box)
```

## Delete Non-ASCII Files

```{r}
files |>
  head(1) |>
  dirname() |>
  dir_ls(
    type = "file",
    regexp = "asc$|md$",
    invert = TRUE
  ) |>
  file_delete()
```
