# Logônia `logolink` Run

```{r}
#| label: Setup
#| include: false

library(here)

here("R", ".setup.R") |> source()
```

## Overview

This document provides a step-by-step guide to run [Logônia](https://github.com/sustentarea/logonia) using the [`logolink`](https://danielvartan.github.io/logolink) R package.

## Methods

### Data Munging

The data munging follows the data science workflow outlined by Wickham et al. ([2023](https://r4ds.hadley.nz/)). All processes were made using the [Quarto](https://quarto.org/) publishing system, the [R](https://www.r-project.org/) programming language, and several R packages.

For data manipulation and workflow, priority was given to packages from the [Tidyverse](https://www.tidyverse.org/) and [rOpenSci](https://ropensci.org/), as well as other packages adhering to the [tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html).

### Code Style

The Tidyverse [tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html), [code style guide](https://style.tidyverse.org/) and [design principles](https://design.tidyverse.org/) were followed to ensure consistency and enhance readability.

### Reproducibility

The pipeline is fully reproducible and can be run again at any time. To ensure consistent results, the [`renv`](https://rstudio.github.io/renv/) package was used to manage and restore the R environment.

## Set Environment

### Load Packages

```{r}
#| label: Set Environment
#| code-fold: false
#| output: false

library(brandr)
library(dplyr)
library(ggplot2)
library(here)
library(janitor)
library(logolink)
library(lubridate)
library(ragg)
library(tidyr)
library(tsibble)
```

### Set Paths

```{r}
model_path <- here("nlogox", "logonia.nlogox")
```

## Run NetLogo Check Experiment

```{r}
#| label: Run NetLogo Check Experiment

results <-
  model_path |>
  run_experiment(
    experiment = "NetLogo Check (default values)"
  )
```

```{r}
results |> glimpse()
```

## Analyze Data

### Prepare Data

```{r}
#| label: Analyze the Data

data <-
  results |>
  extract2("table") |>
  clean_names() |>
  mutate(
    month = start_month |>
      recode_values(
        "January" ~ 1L,
        "February" ~ 2L,
        "March" ~ 3L,
        "April" ~ 4L,
        "May" ~ 5L,
        "June" ~ 6L,
        "July" ~ 7L,
        "August" ~ 8L,
        "September" ~ 9L,
        "October" ~ 10L,
        "November" ~ 11L,
        "December" ~ 12L,
        default = NA_integer_
      ) |>
      add(step) |>
      magrittr::mod(12) %>%
      if_else(. == 0L, 12L, .),
    year = start_year |>
      as.integer() %>%
      add(cumsum(month == 1L) - (first(month) == 1L)),
    year_month = make_yearmonth(year, month),
    across(
      .cols = starts_with("mean_"),
      .fns = ~ if_else(year == start_year, NA_real_, .x)
    )
  ) |>
  select(
    year_month,
    Seedling = count_logonias_with_shape_seedling,
    `Seedling 12 MA` = mean_seedling_12,
    Juvenile = count_logonias_with_shape_juvenile,
    `Juvenile 12 MA` = mean_juvenile_12,
    Adult = count_logonias_with_shape_adult,
    `Adult 12 MA` = mean_adult_12
  ) |>
  pivot_longer(
    cols = c(
      Seedling,
      `Seedling 12 MA`,
      Juvenile,
      `Juvenile 12 MA`,
      Adult,
      `Adult 12 MA`
    ),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(
    type = type |>
      factor(
        levels = c(
          "Seedling",
          "Seedling 12 MA",
          "Juvenile",
          "Juvenile 12 MA",
          "Adult",
          "Adult 12 MA"
        )
      )
  ) |>
  tsibble(
    index = year_month,
    key = type
  )
```

```{r}
data |> glimpse()
```

### Plot Time Series

```{r}
#| warning: false

data |>
  ggplot(
    aes(x = year_month, y = value, color = type)
  ) +
  geom_line() +
  scale_color_manual(
    values = c(
      "Seedling" = "#d5f6d8",
      "Seedling 12 MA" = "#2cd13b",
      "Juvenile" = "#f7f7ad",
      "Juvenile 12 MA" = "#92921d",
      "Adult" = "#ebe2da",
      "Adult 12 MA" = "#9d6e48"
    )
  ) +
  scale_x_yearmonth(
    breaks = data |>
      pull(year_month) |>
      unique() |>
      as.Date() %>%
      if_else(
        month(.) == 1,
        .,
        NA
      ) |>
      na.omit() %>%
      magrittr::extract(seq(1, length(.), by = 10L)) |>
      year() |>
      make_yearmonth(1L)
  ) +
  labs(
    x = "Year-Month",
    y = "Frequency",
    color = NULL
  )
```
