# Logônia Plot Rendering

```{r}
#| label: Setup
#| include: false

library(here)

here("R", ".setup.R") |> source()
```

## Overview

This document provides a step-by-step guide to render the plots used in the  [Logônia](https://github.com/sustentarea/logonia) model documentation.

## Methods

### Source of Data

Data used in this notebook come from the following sources:

- [WorldClim](https://www.worldclim.org/)
  - [Historical Monthly Weather Data](https://www.worldclim.org/data/monthlywth.html) series, including 12 monthly data points for each year from 1951 to 2024. Based on [downscaled](https://worldclim.org/data/downscaling.html) data from [CRU-TS-4.09](https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.09/) (Harris et al., [2020](https://doi.org/10.1038/s41597-020-0453-3)), developed by the [Climatic Research Unit](https://www.uea.ac.uk/groups-and-centres/climatic-research-unit) at the [University of East Anglia](https://www.uea.ac.uk/). It provides monthly averages for minimum temperature, maximum temperature, and total precipitation.
- [GADM](https://gadm.org/): Database of Global Administrative Areas
  - Data on administrative boundaries and regions, utilized for plotting country boundaries and cropping raster datasets.

### Data Munging

The data munging follows the data science workflow outlined by Wickham et al. ([2023](https://r4ds.hadley.nz/)). All processes were made using the [Quarto](https://quarto.org/) publishing system, the [R](https://www.r-project.org/) programming language, and several R packages.

For data manipulation and workflow, priority was given to packages from the [Tidyverse](https://www.tidyverse.org/), [rOpenSci](https://ropensci.org/) and [rspatial](https://rspatial.org/) frameworks, as well as other packages adhering to the [tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html).

### Code Style

The Tidyverse [tidy tools manifesto](https://tidyverse.tidyverse.org/articles/manifesto.html), [code style guide](https://style.tidyverse.org/) and [design principles](https://design.tidyverse.org/) were followed to ensure consistency and enhance readability.

### Reproducibility

The pipeline is fully reproducible and can be run again at any time. To ensure consistent results, the [`renv`](https://rstudio.github.io/renv/) package was used to manage and restore the R environment.

## Set Environment

```{r}
#| label: Set Environment
#| include: false

library(cli)
library(httr2)

if (!is_online()) {
  cli_abort(
    paste0(
      "No {.strong {col_red('internet connection')}} detected. ",
      "Please check your connection and try again."
    )
  )
}
```

### Load Packages

```{r}
#| code-fold: false
#| output: false

library(brandr)
library(checkmate)
library(cli)
library(colorspace)
library(dplyr)
library(fs)
library(geodata)
library(ggplot2)
library(glue)
library(here)
library(httr2)
library(magick)
library(orbis) # github.com/sustentarea/orbis
library(patchwork)
library(ragg)
library(stringr)
library(terra)
```

### Set Data Directories

```{r}
images_dir <- here("images")
```

```{r}
if (!dir_exists(images_dir)) {
  dir_create(images_dir, recurse = TRUE)
}
```

## Plot Sigmoid

### Load Custom Functions

::: {.callout-note}
The source code for the functions below can be found in the [R directory](https://github.com/sustentarea/logonia/tree/main/R) of the code repository.
:::

```{r}
#| label: Plot Sigmoid

here("R", "plot_sigmoid.R") |> source()
```

### Plot Sigmoid

```{r}
plot_sigmoid(
  intercept_range = seq(-235, -197.5, by = 1),
  coefficients = c(tmin = 0.35, tmax = 0.45, prec = 1),
  values = c(tmin = 22.32, tmax = 31.22, prec = 194.61),
  text_size = 8,
  width = 600,
  height = 400,
  file_name = images_dir |> file.path("logistic-regression-sigmoid.png")
)
```

## Plot Animated Data

### Plot Animation 1

#### Load Custom Functions

::: {.callout-note}
The source code for the functions below can be found in the [R directory](https://github.com/sustentarea/logonia/tree/main/R) of the code repository.
:::

```{r}
#| label: Plot Animated Data

here("R", "plot_animated_data.R") |> source()
```

#### Plot Animation

```{r}
plot_animated_data(
  year = 1951,
  months = 3,
  text_size = 18,
  width = 1000,
  height = 300,
  file_name = images_dir |> file.path("worldclim-animation.gif")
)
```

### Plot Animation 2

#### Load Custom Functions

::: {.callout-note}
The source code for the functions below can be found in the [R directory](https://github.com/sustentarea/logonia/tree/main/R) of the code repository.
:::

```{r}
here("R", "plot_animated_data-2.R") |> source()
```

#### Plot Animation

```{r}
plot_animated_data(
  year = 1951,
  months = 3,
  text_size = 18,
  width = 1000,
  height = 300,
  file_name = images_dir |> file.path("worldclim-animation-2.gif")
)
```
